From OpenACC to OpenMP 4: \\ Toward Automatic Translation

### 2016-04-29 paper.tex
Add table with execution times

\documentclass{sig-alternate-05-2015}
\usepackage{xcolor}
\usepackage{xspace}
\hyphenpenalty=750
\newcommand\todo[1]{\textcolor{red}{** #1}}
\newcommand{\pooda}{\textsc{POOD}\xspace}

@@

To evaluate the correctness of our tool, we used it to translate the EPCC
Level~1 OpenACC benchmarks~\cite{epcc}, which are ports of the PolyBench and
PolyBench/GPU kernels~\cite{polybench}.  We also translated several
miscellaneous test programs, including naive and tiled matrix multiplication
kernels.  We compiled the original OpenACC code using PGCC~16.1 and the
resulting OpenMP code using Clang~3.8.  Performance measurements were taken on
a Tesla K40 (Comet and Stampede have Tesla K20 GPUs).

@@

@@

Perhaps more importantly, performance portability is not guaranteed.  A direct
translation of OpenACC code optimized for an NVIDIA GPU will not necessary
perform at its peak on a Xeon Phi.

Interestingly, we did not even obtain comparable performance targeting the
\emph{same} hardware.  In our test cases, the GPU kernels generated by Clang
were almost always several times slower than those generated by PGCC.  For
example, consider the naive matrix multiplication kernel in
Figure~\ref{fig:matmul}.  With $2048 \times 2048$ matrices, the OpenACC kernel
ran for about 250~ms; the OpenMP kernel ran for about 925~ms.  Since the
OpenACC and OpenMP codes were nearly identical, and they were running on the
same hardware, we did not expect such massive differences in performance.
Clearly, details of the compiler-generated code can have a substantial impact
on performance.  Again, documenting differences in compilers would be
helpful---what sort of performance change is typical, what can be done to
improve performance when changing from compiler X to compiler Y, etc.

\begin{figure}
\begin{verbatim}
#pragma omp target teams distribute
#pragma acc parallel loop gang
for (int j = 0; j < N; j++) {
    #pragma omp parallel for
    #pragma acc loop vector
    for (int i = 0; i < N; i++) {
        double t = 0.0;
        #pragma acc loop seq
        for (int k = 0; k < N; k++) {
            t += m[k][i] * n[j][k];
        }
        p[j][i] = t;
    }
}
\end{verbatim}
\caption{Naive matrix multiplication.}
\label{fig:matmul}
\end{figure}

>>>>

\documentclass{sig-alternate-05-2015}
\usepackage[T1]{fontenc}
\usepackage[latin1]{inputenc}
\usepackage{xspace}
\usepackage[table]{xcolor}
\hyphenpenalty=750
\newcommand\todo[1]{\textcolor{red}{** #1}}
\newcommand{\pooda}{\textsc{POOD}\xspace}

@@

To evaluate the correctness of our tool, we used it to translate the EPCC
Level~1 OpenACC benchmarks~\cite{epcc}, which are ports of the PolyBench and
PolyBench/GPU kernels~\cite{polybench}.  We also translated several
miscellaneous test programs, including na\"{i}ve and tiled matrix multiplication
kernels.  We compiled the original OpenACC code using PGCC~16.1 and the
resulting OpenMP code using Clang~3.8.  Performance measurements were taken on
a Tesla K40 (Comet and Stampede have Tesla K20 GPUs).

@@

Interestingly, while both OpenACC and Clang were targeting the same GPU device,
the GPU kernels generated by Clang were almost always slower than those
generated by PGCC.  The naive matrix multiplication kernel in
Figure~\ref{fig:matmul} can serve as a small, concrete example.  With $2048
\times 2048$ matrices, the OpenACC kernel ran for about 250~ms; the OpenMP
kernel ran for about 925~ms.  Since the OpenACC and OpenMP codes were nearly
identical, and they were running on the same hardware, we did not expect such
massive differences in performance.  Execution times for the EPCC Level~1
benchmarks on a Tesla~K40 are shown in Table~\ref{table:epcc}.  The benchmarks
were run with a data size of 1024 and 5 repetitions.  Clearly, details of the
compiler-generated code can have a substantial impact on performance.
%Again, documenting differences in compilers would be helpful---what sort of
%performance change is typical, what can be done to improve performance when
%changing from compiler X to compiler Y, etc.

\begin{figure}
\begin{verbatim}
#pragma omp target teams distribute
#pragma acc parallel loop gang
for (int j = 0; j < N; j++) {
    #pragma omp parallel for
    #pragma acc loop vector
    for (int i = 0; i < N; i++) {
        double t = 0.0;
        #pragma acc loop seq
        for (int k = 0; k < N; k++) {
            t += m[k][i] * n[j][k];
        }
        p[j][i] = t;
    }
}
\end{verbatim}
\caption{Na\"{i}ve matrix multiplication.}
\label{fig:matmul}
\end{figure}

\begin{table}
{\sf
\rowcolors{0}{gray!25}{white}
\begin{tabular}{l@{\hspace*{1em}}r@{\hspace*{1em}}r@{\hspace*{1em}}r}
\textbf{Kernel} & \textbf{PGI/OpenACC} & \textbf{Clang/OpenMP} & \textbf{Speedup} \\
3MM     & 17241 &  50073 & 0.34 \\
ATAX    & 24539 &  28308 & 0.87 \\
BICG    & 25178 &  29561 & 0.85 \\
MVT     & 24021 &  27928 & 0.86 \\
SYRK    & 72562 & 213693 & 0.34  \\
COV     & 28120 &  19905 & 1.41 \\
COR     & 28640 & 187218 & 0.15 \\
SYR2K   & 15231 &  66935 & 0.23 \\
GESUMMV & 37998 &  52765 & 0.72 \\
\multicolumn{4}{r}{\textit{Median: 0.72}}
\end{tabular}
}
\caption{Execution times for EPCC Level~1 benchmarks on a Tesla~K40 ($\mu s$).}
\label{table:epcc}
\end{table}

@@

Perhaps more importantly, performance portability is not guaranteed.  A direct
translation of OpenACC code optimized for an NVIDIA GPU will not necessary
perform at its peak on a Xeon Phi.  As noted above, we did not even obtain
comparable performance targeting the
\emph{same} hardware.
